---
title: "Masticated Fuels Analyses: No Transformation of Response Variable"
author: "Sam Wozniak"
date: "June 6, 2018"
output:
  pdf_document: 
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
library(trtools)
options(digits = 3)
```

```{r, include = FALSE}
d <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\mast_all.csv', header = TRUE)  #Bruce's data
live <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\data_20180513105603.csv', header = TRUE)
tdi <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\tdi_mast.csv', skip = 5, header = TRUE)
tdi <- select(tdi, 1:9)
tdi <- filter(tdi, LOC != 'ST')
```

```{r, include = FALSE}
#remove columns in d that are not useful
columns <- c('X1h', 'X10h', 'X100h', 'X1000h', 'X100h_1000h', 'duff_litter_load', 'tree_cvr_SS')
d <- select(d, -one_of(columns))

# create subplot_id column for d
d$tcode <- 'GC'
d$rcode <- 'JP'
d$pre_sub <- ifelse(d$subplot < 10, paste0('00', d$subplot), paste0('0', d$subplot))
d$pre_sub <- paste0(d$tcode, d$pre_sub)
d$site1 <- d$site
d <- unite(d, subplot_id, rcode, site1, pre_sub, sep = '-') 
```

```{r, include = FALSE}
###combine tdi data with live data.frame
tdi$tcode <- 'GC'
tdi$rcode <- tdi$REG
tdi$scode <- tdi$LOC
tdi$pre_sub <- ifelse(tdi$SUB < 10, paste0('00', tdi$SUB), paste0('0', tdi$SUB))
tdi$pre_sub <- paste0(tdi$tcode, tdi$pre_sub)
tdi <- unite(tdi, subplot_id, rcode, scode, pre_sub, sep = '-') 
minimal <- select(tdi, subplot_id, TDI, TC)

l <- left_join(live, minimal, by = 'subplot_id')
l$site <- l$scode
```

```{r, include = FALSE}
d$dwd_1hr <- d$kgha_1h
d$dwd_10hr <- d$kgha_10h
d$dwd_100_1000hr <- d$kgha_100_1000h
d$pre_tree_cvr <- d$TC

d <- filter(d, !is.na(TC))
```

\pagebreak

#Notes

Years since treatment is treated as a factor. This means that the effect of year since treatment is not an incremental increase (or decrease) from 0 to 1, 1 to 2, or 2 to 3 years since treatment. Instead, there is an effect of year since treatment for each year relative to the reference year (1 yst for downed woody debris, and pre-treatment for all other response variables).

For dead fuels (masticated downed woody debris and tree litter + duff), pre-treatment tree cover is used as an explanatory variable.

For live response variables (e.g. herbaceous fuels, perennial grass cover, shrub cover, tree cover, tree density), pre-treatment Tree Dominance Index (TDI) is used as an explanatory variable. TDI is used instead of pre-treatment tree cover because the response of living plants over time depends upon the relative competition between plant functional groups at the time of treatment.

#Definitions of acronyms and abbreviations used in code

yst = years since treatment; 0 represents pre-treatment

TDI = Tree Dominance Index (pre-treatment tree cover/ (pre-treatment perennial grass cover + pre-t. shrub cover + pre-t. tree cover))

pre_tree_cvr = pre-treatment tree cover (%)

dwd_1hr = Downed woody debris of 1-hr class (< 1/4 in diameter)

dwd_10hr = Downed woody debris of 10-hr class (1/4 - 1 in diameter)

dwd_100_1000hr = Downed woody debris of 100-hr and 1000-hr classes (1-3 and 3+ inches diameter); these two fuel classes were combined because 1000-hr fuels are very infrequent (vast majority of subplots have 0 1000-hr fuels) on masticated sites 

herb_ttl = live + dead herbaceous fuel loading

\pagebreak

#Masticated 1-hr fuels

##Notes

Sampling at 5-6 years: 2/3 sites (GR, SC) were sampled 5 years since treatment, and 1/3 sites (ON) was sampled at 6 years post-treatment

##Model
```{r}
m <- lmer(dwd_1hr ~ pre_tree_cvr + yst + pre_tree_cvr:yst + 
            (1|site), data = d)
```



```{r, echo = FALSE}
#Response averaged across sites
d$yhat1 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_1hr))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat1))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 1-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + facet_wrap(~yst, ncol = 2)
plot(p)
```

\pagebreak

##Inferences 
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + 
  facet_wrap(~yst, ncol = 2) + labs(title = 'Residuals by Years Since Treatment')
```


```{r, eval = FALSE, include = FALSE}
#by yst and site
d$yhat1 <- predict(m)

p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_1hr, color = site))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat1))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 1-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + facet_wrap(site~yst)
plot(p)


ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + 
  facet_wrap(site~yst) + labs(title = 'Residuals by Years Since Treatment and Site')
```

\pagebreak

#Masticated 10-hr fuels

##Model
```{r}
m <- lmer(dwd_10hr ~ pre_tree_cvr + yst + pre_tree_cvr:yst + 
            (1|site), data = d)
```



```{r, echo = FALSE}
#Response averaged across sites
d$yhat10 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_10hr))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat10))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 10-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + facet_wrap(~yst)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(~yst, ncol = 2) + 
  labs(title = 'Residuals by Years Since Treatment')
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
d$yhat10 <- predict(m)
p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_10hr, color = site))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat10))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 10-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + facet_wrap(site~yst)
plot(p)

ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(site~yst) + 
  labs(title = 'Residuals by Years Since Treatment and Site')
```

\pagebreak

#Masticated 100 + 1000-hr fuels


##Model
```{r}
m <- lmer(dwd_100_1000hr ~ pre_tree_cvr + factor(yst) + pre_tree_cvr:factor(yst) + 
            (1|site), data = d)
```



```{r, echo = FALSE}
#Response avereaged across sites
d$yhat100_1000 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_100_1000hr))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat100_1000))
p <- p + theme_classic()
p <- p + facet_wrap(~yst, ncol = 2)
p <- p + labs(title = 'Masticated 100-hr + 1000-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(~yst, ncol = 2) + 
  labs(title = 'Residuals by Years Since Treatment')
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
d$yhat100_1000 <- predict(m)

p <- ggplot(data = d, aes(x = pre_tree_cvr, y = dwd_100_1000hr, color = site))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat100_1000))
p <- p + theme_classic()
p <- p + facet_wrap(site~yst)
p <- p + labs(title = 'Masticated 100-hr + 1000-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)', 
                color = 'site')
plot(p)

ggplot(m, aes(x = .fitted, y = .resid, color = site)) + 
  geom_point() + 
  facet_wrap(site~yst) + 
  labs(title = 'Residuals by Years Since Treatment and Site',
       color = 'site')
```

\pagebreak

#Tree Litter + Duff Fuels

##Notes

In model code, 'duff' refers to tree litter + duff.

```{r, include = FALSE}
data <- read.csv("C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\duff.csv",
                 header = TRUE) 

#remove observations where year = NA
data <- data[-which(is.na(data$year)),]

#remove observations where sp_phase = NA
data <- data[-which(is.na(data$sp_phase)),]

#remove observation where duff load = Infinity
data <- data[-which(data$tree_ltr_duff_ld > 1000000),]

#keep only the columns/variables of interest
data <- data[, c("subplot_id", "year", "tree_ltr_ld", "tree_duff_ld", 
                 "tree_cover_ttl", "tree_ltr_duff_ld", "scode", 
                 "rcode", "treatment", "sp_phase")]

#create relational table with site code and implementation year
scode <- c('ON', 'BC', 'WB', 'MC','BM', 'DR', 'SV', 'GR', 'SC', 'SR')
y_imp <- c(6, 6, 6, 6, 7, 7, 7, 7, 7, 8)
relation <- data.frame(scode, y_imp)

d <- left_join(data, relation, by = "scode")

#create year since treatment column
d$yst <- d$year - d$y_imp

#create column with pre-treatment tree cover titled 'pre_tree_cvr'
d$pre_tree_cvr <- NA
for (i in unique(d$subplot_id)) {                                                            #for each unique subplot_id
  pre_tree_c <- d$tree_cover_ttl[d$subplot_id == i][which.min(d$year)]                      #take the tree cover value from the first 
  d$pre_tree_cvr[d$subplot_id == i] <- pre_tree_c                                                #put those tree cover values in a column
}

#subset columns for yst <=1 or yst = 10
a <- d[which(d$yst == 1 | d$yst == 10),]
b <- a[!(a$scode == 'SR'),]

#subset rows where yst = 1 and tree_ltr_ld = NA and tree_duff_ld = NA
na0 <- subset(b, !(yst == 1 & is.na(tree_ltr_ld) & is.na(tree_duff_ld)))
b <- b[-which(b$yst == 1 & is.na(b$tree_ltr_ld) & is.na(b$tree_duff_ld)),]

#create column with tree litter + duff load for every year
b$duff <- ifelse(b$yst == 10, b$tree_ltr_duff_ld, rowSums(b[,c("tree_ltr_ld", 
                                                               "tree_duff_ld")], 
                                                          na.rm=TRUE))
b[which(is.na(b$duff)),]
b <- b[-which(is.na(b$duff)),]

d <- filter(b, treatment == 'BM')

d$site <- d$scode

```

##Model
```{r}
m <- lmer(duff ~ factor(yst) + pre_tree_cvr + factor(yst):pre_tree_cvr + 
            (1|site), data = d)
```


```{r, echo = FALSE}
d$yhat_duff <- predict(m, re.form = NA)
p <- ggplot(data = d, aes(x = pre_tree_cvr, y = duff))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_duff))
p <- p + theme_classic() + facet_wrap(~yst, ncol = 2)
p <- p + labs(title = 'Tree Litter + Duff',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
plot(p)
```

\pagebreak

##Inferences 
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
d$yhat_duff <- predict(m)
p <- ggplot(data = d, aes(x = pre_tree_cvr, y = duff, color = site))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_duff))
p <- p + theme_classic() + facet_wrap(site~yst, ncol = 2)
p <- p + labs(title = 'Tree Litter + Duff',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'Year Since \nTreatment')
plot(p)

ggplot(m, aes(x = .fitted, y = .resid, color = site)) + 
  geom_point() + facet_wrap(site~yst, nrow = 3)
```

\pagebreak

#Herbaceous fuel loading (live + dead)

##Notes:
*Investigate value of zero at Onaqui, yst = 10

```{r, echo = FALSE}
scode <- c('ON', 'GR', 'SC')
y_imp <- c(6, 7, 7)
relation <- data.frame(scode, y_imp)

l <- left_join(l, relation, by = "scode")

l$herb_ttl <- l$herb_fuel_dead_wd + l$herb_fuel_live_wd
l <- filter(l, !is.na(herb_ttl))

l$yst <- l$year - l$y_imp

td <- l #for tree density and tree cover later on

l <- filter(l, yst %in% c(0,1,2,3,6,10))
```

```{r, include = FALSE}
precipitation <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\Bruce\\edited\\precip_b_and_prism.csv', 
                          header = TRUE)
precip <- precipitation %>% 
  select(scode, yst, OJprecip, source) %>%
  filter(yst %in% c(0,1,2,3,6,10))
l <- left_join(l, precip, by = c('scode', 'yst'))
l$site <- l$scode

#m <- lmer(sqrt(herb_ttl) ~ TDI + yst + yst:TDI + (1|scode) + (1|OJprecip), data = l)
```

##Model
```{r}
m <- lmer(herb_ttl ~ TDI + factor(yst) + factor(yst):TDI + 
            (1|site), data = l)
```

```{r, echo = FALSE}
#by yst
l$yhat_herb <- predict(m, re.form = NA)
p <- ggplot(data = l, aes(x = TDI, y = herb_ttl))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_herb))
p <- p + theme_bw()
p <- p + labs(title = 'Herbaceous Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
l$yhat_herb <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = herb_ttl, color = site))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_herb))
p <- p + theme_bw()
p <- p + labs(title = 'Herbaceous Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(site~yst, ncol = 6, nrow = 3)
plot(p)


ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(site~yst, nrow = 3, ncol = 6)
```

\pagebreak

#Shrub Fuels
*Data error: two values of zero at Years since treatment = 6 & TDI ~ 0.3; zero values are incorrect (JP-ON-GC-006, JP-ON-GC-010 have high shrub volumes but zero biomass)

*Missing data: no shrub data for Onaqui when YST = 0 (calendar year = 2006)

```{r, echo = FALSE}
l$shrub_fuel <- abs(l$shrub_bio_ttl)

l$shrub_fuel[l$scode == 'ON' & l$yst == 0] <- NA

l2 <- subset(l, !is.na(shrub_fuel))
l2$scode <- factor(l2$scode, levels = c('GR', 'SC', 'ON'))
l2$site <- l2$scode
```

##Model
```{r}
m <- lmer(shrub_fuel ~ TDI + factor(yst) + factor(yst):TDI + 
            (1|site), data = l2)
```


```{r, echo = FALSE}
#by yst
l2$yhat_shrub <- predict(m, re.form = NA)
p <- ggplot(data = l2, aes(x = TDI, y = shrub_fuel))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_shrub))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = .3))
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
l2$yhat_shrub <- predict(m)
p <- ggplot(data = l2, aes(x = TDI, y = shrub_fuel, color = site))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_shrub))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = .3))
p <- p + facet_wrap(site ~ yst, ncol = 6, nrow = 3)
plot(p)

ggplot(m, aes(x = .fitted, y = .resid, color = site)) + 
  geom_point() + 
  facet_wrap(site~yst, ncol = 6, nrow = 3)
```

\pagebreak

#Shrub Cover

##Notes

Shrub cover increase when yst = 6 for site = SC & GR but decrease in herb biomass

##Model
```{r}
m <- lmer(can_cover_pt_shrub ~ TDI + factor(yst) + factor(yst):TDI + 
            (1|site), data = l)
```



```{r, echo = FALSE}
#by yst
l$yhat_sh_cvr <- predict(m, re.form = NA)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_shrub))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_sh_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Cover',
                x = 'Tree Dominance Index',
                y = 'Shrub Cover (%)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```



```{r, eval = FALSE, include = FALSE}
#by yst and site
l$yhat_sh_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_shrub, color = site))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_sh_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Cover',
                x = 'Tree Dominance Index',
                y = 'Fuel Loading (kg/ha)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(site~yst, ncol = 6, nrow = 3)
plot(p)


ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(site~yst, ncol = 6, nrow = 3)
```

\pagebreak

#Perennial Grass Cover

##Model
```{r}
m <- lmer(can_cover_pt_pgrass ~ TDI + factor(yst) + factor(yst):TDI + 
            (1|site), data = l)
```


```{r, echo = FALSE}
#by yst
l$yhat_pgrass_cvr <- predict(m, re.form = NA)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_pgrass))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_pgrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Perennial Tall Grass Cover',
                x = 'Pre-treatment Tree Dominance Index',
                y = 'Square Root of Perennial Tall Grass Cover (%)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

```{r, eval = FALSE, include = FALSE}
#by yst and site
l$yhat_pgrass_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_pgrass, color = site))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_pgrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Perennial Tall Grass Cover',
                x = 'Tree Dominance Index',
                y = 'Perennial Tall Grass Cover (%)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(site~yst, ncol = 6, nrow = 3)
plot(p)

ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(site~yst, ncol = 6, nrow = 3)
```

\pagebreak

#Annual Grass Cover

##Notes

what is going on at Scipio in yst = 6,10? Decrease in annual grass cover

##Model
```{r}
m <- lmer(can_cover_pt_agrass ~ TDI + factor(yst) + factor(yst):TDI + 
            (1|site), data = l)
```



```{r, echo = FALSE}
#by yst
l$yhat_agrass_cvr <- predict(m, re.form = NA)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_agrass))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_agrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Annual Grass Cover',
                x = 'Pre-treatment Tree Dominance Index',
                y = 'Square Root of Annual Grass Cover (%)')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences 
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

```{r, include = FALSE}
#by yst and site
l$yhat_agrass_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_agrass, color = site))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_agrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Annual Grass Cover',
                x = 'Tree Dominance Index',
                y = 'Square Root of Annual Grass Cover (%)', 
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(site~yst, ncol = 6, nrow = 3)
plot(p)

ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(site~yst, ncol = 6, nrow = 3)
```

\pagebreak

#Notes on Tree Density and Cover

*Ask Scott--was tree density for trees > 50 cm measured at 1,2,3,6 yst?

*Should I break this down by species (JUOS vs PIED)?

```{r, echo = FALSE}
#td$tree_density <- td$tree_dns_5_50_JUOS + td$tree_dns_gt50_JUOS + td$tree_dns_5_50_PIED + td$tree_dns_gt50_PIED
td <- filter(td, (yst %in% c(-1,1,2,3,6,10) & scode %in% c('SC', 'GR')) | 
               (yst %in% c(0,1,2,3,6,10) & scode == 'ON'))
td$yst[td$yst == -1] <- 0  #so that all pre-treatment years are grouped together

td$site <- td$scode
```

#Tree Density for trees < 5 cm in height

##Notes

*Model and graph below are included to show that tree density for trees < 5cm should be excluded from analysis. Data is too coarse: if there was one tree < 5 cm found in a subplot, that converts to 889 trees/ha because of sampling density and scaling factor. 

##Model
```{r, include = FALSE}
td$tree_dns_lt5 <- td$tree_dns_lt5_JUOS + td$tree_dns_lt5_PIED + td$tree_dns_lt5_PIMO
```

```{r}
m <- lmer(tree_dns_lt5 ~ TDI + factor(yst) + factor(yst):TDI + (1|site), 
          data = td)
```

```{r, echo = FALSE}
#by yst
td$yhat_tree_dns_lt5 <- predict(m, re.form = NA)
p <- ggplot(data = td, 
            aes(x = TDI, y = tree_dns_lt5))
p <- p + geom_point()
p <- p + theme_bw()
p <- p + geom_line(aes(y = yhat_tree_dns_lt5))
p <- p + labs(title = 'Tree Density for trees < 5 cm in height',
                x = 'Pre-treatment Tree Dominance Index',
                y = 'Square Root of Tree Density (stems/ha)')
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)

hist(td$tree_dns_lt5)
```

\pagebreak

#Tree Density for trees between 5 and 50 cm in height

##Model

```{r, include = FALSE}
td$tree_dns_5_50 <- td$tree_dns_5_50_JUOS + td$tree_dns_5_50_PIED + td$tree_dns_5_50_PIMO
```


```{r}
m <- lmer(tree_dns_5_50 ~ TDI + factor(yst) + factor(yst):TDI + (1|site), 
          data = td)
```

```{r, echo = FALSE}
#by yst
td$yhat_tree_dns_5_50 <- predict(m, re.form = NA)
p <- ggplot(data = td, 
            aes(x = TDI, y = tree_dns_5_50))
p <- p + geom_point()
p <- p + theme_bw()
p <- p + geom_line(aes(y = yhat_tree_dns_5_50))
p <- p + labs(title = 'Tree Density for trees between 5 and 50 cm in height',
                x = 'Pre-treatment Tree Dominance Index (%)',
                y = 'Square Root of Tree Density (stems/ha)')
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

\pagebreak

#Tree Density for trees > 50 cm in height

##Notes

*Ask Scott--was tree density for trees > 50 cm measured at 1,2,3,6 yst?

```{r, echo = FALSE}
td$tree_dns_gt50 <- td$tree_dns_gt50_JUOS + td$tree_dns_gt50_PIED + td$tree_dns_gt50_PIMO
```

##Model
```{r}
m <- lmer(tree_dns_gt50 ~ TDI + factor(yst) + factor(yst):TDI + (1|site), 
          data = td)
```


```{r, echo = FALSE}
#by yst
td$yhat_tree_dns_gt50 <- predict(m, re.form = NA)
p <- ggplot(data = td, 
            aes(x = TDI, y = tree_dns_gt50))
p <- p + geom_point()
p <- p + theme_bw()
p <- p + geom_line(aes(y = yhat_tree_dns_gt50))
p <- p + labs(title = 'Tree Density for trees > 50 cm in height',
                x = 'Pre-Treatment Tree Dominance Index (%)',
                y = 'Square Root of Tree Density (stems/ha)')
p <- p + facet_wrap(~yst, ncol = 3)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r, echo = FALSE}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
```

\pagebreak

#Tree Cover 

##Notes

*Method: measured canopy area of trees > 50 cm in height and divided by area of subplot

```{r, echo = FALSE}
#Pre-treatment tree cover for all sites was collected in 2006
tcover <- filter(td, site == 'ON' & year %in% c(6,16)|
                   site == 'GR' & year %in% c(6,17)|
                   site == 'SC' & year %in% c(6,17))
tcover$tree_cover_ttl[tcover$subplot_id %in% 
                        c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & 
                        tcover$yst == 10] <- 0
tcover$tree_cvr_PIED[tcover$subplot_id %in% 
                       c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & 
                       tcover$yst == 10] <- 0
tcover$tree_cvr_JUOS[tcover$subplot_id %in% 
                       c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & 
                       tcover$yst == 10] <- 0

#tcover$yst <- factor(tcover$yst, levels = c(-1,0,10))
#levels(tcover$yst) <- c('pre', 'pre', '10')
```

##Model

```{r}
m <- lmer(tree_cover_ttl ~ TDI + factor(yst) + factor(yst):TDI + (1|site), 
          data = tcover)
```

```{r, echo = FALSE}
#by yst
tcover$yhat_tree_cover <- predict(m, re.form = NA)
p <- ggplot(data = tcover, aes(x = TDI, y = tree_cover_ttl))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_tree_cover))
p <- p + theme_bw()
p <- p + labs(title = 'Tree Cover',
                x = 'Pre-treatment Tree Dominance Index',
                y = 'Square Root of Tree Cover (%)')

p <- p + facet_wrap(~yst, ncol = 2)
plot(p)
```

\pagebreak

##Inferences
```{r, echo = FALSE}
summary(m)
lincon(m)
```

\pagebreak

##QQPlot and Plotted Residuals
```{r}
qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + 
  facet_wrap(~yst) 
```

```{r, eval = FALSE, include = FALSE}
tcover$yhat_tree_cover <- predict(m)
p <- ggplot(data = tcover, aes(x = TDI, y = tree_cover_ttl, color = site))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_tree_cover))
p <- p + theme_bw()
p <- p + labs(title = 'Tree Cover',
                x = 'Pre-Treatment Tree Cover (%)',
                y = 'Square Root of Tree Cover (%)')
#p <- p + scale_x_continuous(breaks = seq(0,60, by = 10))
p <- p + facet_wrap(site~yst, ncol = 2, nrow = 3)
plot(p)
```