---
title: "Masticated Fuels Analyses"
author: "Sam Wozniak"
date: "May 13, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
library(trtools)
options(digits = 3)
```

```{r, include = FALSE}
d <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\mast_all.csv', header = TRUE)  #Bruce's data
live <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\data_20180513105603.csv')
tdi <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\tdi_mast.csv', skip = 5, header = TRUE)
tdi <- select(tdi, 1:9)
tdi <- filter(tdi, LOC != 'ST')
```

```{r, include = FALSE}
#remove columns in d that are not useful
columns <- c('X1h', 'X10h', 'X100h', 'X1000h', 'X100h_1000h', 'duff_litter_load', 'tree_cvr_SS')
d <- select(d, -one_of(columns))

# create subplot_id column for d
d$tcode <- 'GC'
d$rcode <- 'JP'
d$scode <- d$site
d$pre_sub <- ifelse(d$subplot < 10, paste0('00', d$subplot), paste0('0', d$subplot))
d$pre_sub <- paste0(d$tcode, d$pre_sub)
d <- unite(d, subplot_id, rcode, site, pre_sub, sep = '-') 
```

```{r, include = FALSE}

###combine tdi data with live data.frame
tdi$tcode <- 'GC'
tdi$rcode <- tdi$REG
tdi$scode <- tdi$LOC
tdi$pre_sub <- ifelse(tdi$SUB < 10, paste0('00', tdi$SUB), paste0('0', tdi$SUB))
tdi$pre_sub <- paste0(tdi$tcode, tdi$pre_sub)
tdi <- unite(tdi, subplot_id, rcode, scode, pre_sub, sep = '-') 
minimal <- select(tdi, subplot_id, TDI, TC)

l <- left_join(live, minimal, by = 'subplot_id')
```

```{r, include = FALSE}
#make year since treatment a factor
#d$yst <- factor(d$yst)
#levels(d$yst) <- c('1', '5-6', '10')

d <- filter(d, !is.na(sp_phase))
```

#Masticated 1-hr fuels
TC = pre-treatment tree cover (%)

yst = years since treatment; 0 represents pre-treatment; yst is a factor for masticated fuels

scode = site

```{r}
m <- lmer(sqrt(kgha_1h) ~ TC + yst + TC:yst + (1 + yst|scode), data = d)
summary(m)
lincon(m)

#by yst; averaged across scode (sites)
d$yhat1 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_1h)))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat1))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 1-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + facet_wrap(~yst, ncol = 4)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + 
  facet_wrap(~yst, ncol = 4) + labs(title = 'Residuals by Years Since Treatment')

d$yhat1 <- predict(m)

#by yst and scode
p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_1h), color = scode))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat1))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 1-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + facet_wrap(scode~yst)
plot(p)


ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + 
  facet_wrap(scode~yst) + labs(title = 'Residuals by Years Since Treatment and Site')
```

#Masticated 10-hr fuels

TC = pre-treatment tree cover (%)

yst = years since treatment
```{r}
m <- lmer(sqrt(kgha_10h) ~ TC + yst + TC:yst + (1 + yst|scode), data = d)
summary(m)
lincon(m)

#by yst; averaged across scode (sites)
d$yhat10 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_10h)))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat10))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 10-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
p <- p + facet_wrap(~yst)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(~yst, ncol = 4) + 
  labs(title = 'Residuals by Years Since Treatment')

#by yst and scode
d$yhat10 <- predict(m)
p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_10h), color = scode))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat10))
p <- p + theme_classic()
p <- p + labs(title = 'Masticated 10-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + facet_wrap(scode~yst)
plot(p)

ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst) + 
  labs(title = 'Residuals by Years Since Treatment and Site')

```

#Masticated 100 + 1000-hr fuels

Need to check zero values for year 1

```{r}
m <- lmer(sqrt(kgha_100_1000h) ~ TC + yst + TC:yst + (1 + yst|scode), data = d)
summary(m)
lincon(m)

#by yst; averaged across scode (sites)
d$yhat100_1000 <- predict(m, re.form = NA)

p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_100_1000h)))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat100_1000))
p <- p + theme_classic()
p <- p + facet_wrap(~yst, ncol = 4)
p <- p + labs(title = 'Masticated 100-hr + 1000-hr Fuels by Years Since Treatment',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)')
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(~yst, ncol = 4) + 
  labs(title = 'Residuals by Years Since Treatment')

#by yst and scode
d$yhat100_1000 <- predict(m)

p <- ggplot(data = d, aes(x = TC, y = sqrt(kgha_100_1000h), color = scode))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat100_1000))
p <- p + theme_classic()
p <- p + facet_wrap(scode~yst)
p <- p + labs(title = 'Masticated 100-hr + 1000-hr Fuels by Years Since Treatment and Site',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)', 
                color = 'site')
plot(p)

ggplot(m, aes(x = .fitted, y = .resid, color = scode)) + 
  geom_point() + 
  facet_wrap(scode~yst) + 
  labs(title = 'Residuals by Years Since Treatment and Site',
       color = 'site')
```

#Tree Litter + Duff Fuels
```{r, include = FALSE}
data <- read.csv("C:\\Users\\User\\Documents\\GitHub\\thesis\\mastication_analysis\\duff.csv",
                 header = TRUE) 

#remove observations where year = NA
data <- data[-which(is.na(data$year)),]

#remove observations where sp_phase = NA
data <- data[-which(is.na(data$sp_phase)),]

#remove observation where duff load = Infinity
data <- data[-which(data$tree_ltr_duff_ld > 1000000),]

#keep only the columns/variables of interest
data <- data[, c("subplot_id", "year", "tree_ltr_ld", "tree_duff_ld", 
                 "tree_cover_ttl", "tree_ltr_duff_ld", "scode", 
                 "rcode", "treatment", "sp_phase")]

#create relational table with site code and implementation year
scode <- c('ON', 'BC', 'WB', 'MC','BM', 'DR', 'SV', 'GR', 'SC', 'SR')
y_imp <- c(6, 6, 6, 6, 7, 7, 7, 7, 7, 8)
relation <- data.frame(scode, y_imp)

d <- left_join(data, relation, by = "scode")

#create year since treatment column
d$yst <- d$year - d$y_imp

#create column with pre-treatment tree cover titled 'pre_tc'
d$pre_tc <- NA
for (i in unique(d$subplot_id)) {                                                            #for each unique subplot_id
  pre_tree_c <- d$tree_cover_ttl[d$subplot_id == i][which.min(d$year)]                      #take the tree cover value from the first 
  d$pre_tc[d$subplot_id == i] <- pre_tree_c                                                #put those tree cover values in a column
}

#subset columns for yst <=1 or yst = 10
a <- d[which(d$yst == 1 | d$yst == 10),]
b <- a[!(a$scode == 'SR'),]

#subset rows where yst = 1 and tree_ltr_ld = NA and tree_duff_ld = NA
na0 <- subset(b, !(yst == 1 & is.na(tree_ltr_ld) & is.na(tree_duff_ld)))
b <- b[-which(b$yst == 1 & is.na(b$tree_ltr_ld) & is.na(b$tree_duff_ld)),]

#create column with tree litter + duff load for every year
b$duff <- ifelse(b$yst == 10, b$tree_ltr_duff_ld, rowSums(b[,c("tree_ltr_ld", 
                                                               "tree_duff_ld")], 
                                                          na.rm=TRUE))
b[which(is.na(b$duff)),]
b <- b[-which(is.na(b$duff)),]

d <- filter(b, treatment == 'BM')
#d$yst <- factor(d$yst)
```

```{r}
#model, inferences, and residuals
m <- lmer(sqrt(duff) ~ yst + pre_tc + yst:pre_tc + (1 + yst|scode), data = d)
summary(m)
lincon(m)

d$yhat_duff <- predict(m)
p <- ggplot(data = d, aes(x = pre_tc, y = sqrt(duff), color = scode))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_duff))
p <- p + theme_classic() + facet_wrap(scode~yst, ncol = 2)
p <- p + labs(title = 'Tree Litter + Duff',
                x = 'Pre-treatment Tree Cover (%)',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'Year Since \nTreatment')
plot(p)

qqnorm(resid(m)); qqline(resid(m))

ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)

ggplot(m, aes(x = .fitted, y = .resid, color = scode)) + 
  geom_point() + facet_wrap(scode~yst, nrow = 3)
```

#For shrub and Herbaceous biomass and cover, use tree dominance index (TDI) instead of pre-treatment tree cover.

Tree Dominance Index (TDI) = (pre-treatment tree cover)/(pre-treatment tree cover + grass cover + shrub cover)

#Herbaceous Fuels
yst = years since treatment

scode = site

herb_ttl = herbaceous fuel loading

**Investigate value of zero at Onaqui, yst = 10
```{r, include = FALSE}
scode <- c('ON', 'GR', 'SC')
y_imp <- c(6, 7, 7)
relation <- data.frame(scode, y_imp)

l <- left_join(l, relation, by = "scode")

l$herb_ttl <- l$herb_fuel_dead_wd + l$herb_fuel_live_wd
l <- filter(l, !is.na(herb_ttl))

l$yst <- l$year - l$y_imp

td <- l #for tree density and tree cover later on

l <- filter(l, yst %in% c(0,1,2,3,6,10))
```

```{r}
precipitation <- read.csv('C:\\Users\\User\\Documents\\GitHub\\thesis\\Bruce\\edited\\precip_b_and_prism.csv', 
                          header = TRUE)
precip <- precipitation %>% 
  select(scode, yst, OJprecip, source) %>%
  filter(yst %in% c(0,1,2,3,6,10))
l <- left_join(l, precip, by = c('scode', 'yst'))
```

```{r}
#m <- lmer(sqrt(herb_ttl) ~ TDI + yst + yst:TDI + (scode|yst), data = l)
m <- lmer(herb_ttl ~ TDI + yst + yst:TDI + (1 + yst|scode) + (1|OJprecip), data = l)
summary(m)
lincon(m)


l$yhat_herb <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = herb_ttl, color = scode))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_herb))
p <- p + theme_bw()
p <- p + labs(title = 'Herbaceous Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst, nrow = 3, ncol = 6)
```

#Shrub Fuels
**Data errors: Onaqui year 6, zero values are incorrect (JP-ON-GC-006, JP-ON-GC-010 have high shrub volumes but zero biomass)

**Missing data: no shrub data for Onaqui when YST = 0 (calendar year = 2006)

shrub_fuel = shrub fuel loading 

TDI = tree dominance index

yst = years since treatment

scode = site
```{r}
l$shrub_fuel <- abs(l$shrub_bio_ttl)

m <- lmer(sqrt(shrub_fuel) ~ TDI + yst + yst:TDI + (1 + yst|scode), data = l)
summary(m)
lincon(m)


l$yhat_shrub <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = sqrt(shrub_fuel), color = scode))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_shrub))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Fuels',
                x = 'Tree Dominance Index',
                y = 'Square Root of Fuel Loading (kg/ha)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = .3))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst, ncol = 6, nrow = 3)
```

###Shrub Cover

Note: Shrub cover increase when yst = 6 for scode = SC & GR but decrease in herb biomass
```{r}
m <- lmer(can_cover_pt_shrub ~ TDI + yst + yst:TDI + (1 + yst|scode), data = l)
summary(m)
lincon(m)


l$yhat_sh_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = can_cover_pt_shrub, color = scode))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_sh_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Shrub Cover',
                x = 'Tree Dominance Index',
                y = 'Fuel Loading (kg/ha)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst, ncol = 6, nrow = 3)
```

###Perennial Grass Cover
```{r}
m <- lmer(sqrt(can_cover_pt_pgrass) ~ TDI + yst + yst:TDI + (1 + yst|scode), data = l)
summary(m)
lincon(m)


l$yhat_pgrass_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = sqrt(can_cover_pt_pgrass), color = scode))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_pgrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Perennial Tall Grass Cover',
                x = 'Tree Dominance Index',
                y = 'Perennial Tall Grass Cover (%)',
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst, ncol = 6, nrow = 3)
```

###Annual Grass Cover
Note: what is going on at Scipio in yst = 6,10? Decrease in annual grass cover
```{r}
m <- lmer(sqrt(can_cover_pt_agrass) ~ TDI + yst + yst:TDI + (1|scode), data = l)
summary(m)
lincon(m)


l$yhat_agrass_cvr <- predict(m)
p <- ggplot(data = l, aes(x = TDI, y = sqrt(can_cover_pt_agrass), color = scode))
p <- p + geom_point()
p <- p + geom_line(aes(y = yhat_agrass_cvr))
p <- p + theme_bw()
p <- p + labs(title = 'Annual Grass Cover',
                x = 'Tree Dominance Index',
                y = 'Square Root of Annual Grass Cover (%)', 
                color = 'site')
p <- p + scale_x_continuous(breaks = seq(0,1, by = 0.3))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point() + 
  facet_wrap(scode~yst, ncol = 6, nrow = 3)
```

###Tree Density >5 cm

```{r}
td$tree_density <- td$tree_dns_5_50_JUOS + td$tree_dns_gt50_JUOS + td$tree_dns_5_50_PIED + td$tree_dns_gt50_PIED
td <- filter(td, (yst %in% c(-1,1,2,3,6,10) & scode %in% c('SC', 'GR')) | 
               (yst %in% c(0,1,2,3,6,10) & scode == 'ON'))
td$yst[td$yst == -1] <- 0  #so that all pre-treatment years are grouped together
```

MODEL FAILS TO CONVERGE UNLESS I TREAT YST AS FACTOR
```{r}
m <- lmer(sqrt(tree_dns_gt50_JUOS + tree_dns_gt50_PIED) ~ TC + factor(yst) + factor(yst):TC + (1 + factor(yst)|scode), data = td)
summary(m)
lincon(m)


td$yhat_tree_dens <- predict(m)
p <- ggplot(data = td, aes(x = TC, y = sqrt(tree_dns_gt50_JUOS + tree_dns_gt50_PIED), color = scode))
p <- p + geom_point()
p <- p + theme_bw()
p <- p + geom_line(aes(y = yhat_tree_dens))
p <- p + labs(title = 'Tree Density for trees > 50 cm',
                x = 'Pre-Treatment Tree Cover (%)',
                y = 'Square Root of Tree Density (stems/ha)',
                color = 'site')
#p <- p + scale_x_continuous(breaks = seq(0,10, by = 2))
p <- p + facet_wrap(scode~yst, ncol = 6, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(scode~yst, nrow = 3)
```

###Tree Cover (trees > 50cm)

```{r, eval = FALSE}
tcover <- filter(td, scode == 'ON' & year %in% c(6,16)|scode == 'GR' & year %in% c(6,17)|scode == 'SC' & year %in% c(6,17))
tcover$tree_cover_ttl[tcover$subplot_id %in% c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & tcover$yst == 10] <- 0
tcover$tree_cvr_PIED[tcover$subplot_id %in% c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & tcover$yst == 10] <- 0
tcover$tree_cvr_JUOS[tcover$subplot_id %in% c('JP-SC-GC002', 'JP-SC-GC004', 'JP-SC-GC007') & tcover$yst == 10] <- 0

m <- lmer(sqrt(tree_cover_ttl) ~ TC + yst + yst:TC + (1 + yst|scode), data = tcover)
summary(m)
lincon(m)


tcover$yhat_tree_cover <- predict(m)
p <- ggplot(data = tcover, aes(x = TC, y = sqrt(tree_cover_ttl), color = scode))
p <- p + geom_jitter()
p <- p + geom_line(aes(y = yhat_tree_cover))
p <- p + theme_bw()
p <- p + labs(title = 'Tree Cover',
                x = 'Pre-Treatment Tree Cover (%)',
                y = 'Square Root of Tree Cover (%)')
#p <- p + scale_x_continuous(breaks = seq(0,60, by = 10))
p <- p + facet_wrap(scode~yst, ncol = 2, nrow = 3)
plot(p)

qqnorm(resid(m)); qqline(resid(m))
ggplot(m, aes(x = .fitted, y = .resid)) + geom_point() + facet_wrap(~yst)
ggplot(m, aes(x = .fitted, y = .resid)) + 
  geom_point(aes(color = scode)) + 
  
  facet_wrap(scode~yst, ncol = 2)
```

